### 2011-09-28 doc/manuscript.tex
The result and UCSC genome browser resource is more refined to do something about virulence genes.

Inferring homologous recombination and horizontal gene transfer 
using whole genomes of \emph{Streptococcus pyogenes} 
and\emph{ S.\ dysgalactiae} subspecies \emph{equisimilis}

@@

phosphotransferase system (PTS), ATP-binding cassette (ABC), Aminoacyl
tRNA synthetase (aaRS), ancestral recombination graph (ARG), group
A streptococci (GAS), group C streptococci (GCS), group G streptococci
(GGS), \emph{Streptococcus pyogenes} (SPY), \emph{S.\ dysgalactiae}
subspecies \emph{dysgalactiae} (SDD), \emph{S.\ dysgalactiae} subspecies
\emph{equisimilis} (SDE).

@@

\section*{Abstract}
Homologous recombination and horizontal gene transfer are investigated
with complete genomes of closely related bacterial species of 
\emph{Streptococcus pyogenes} (SPY) 
and \emph{S.\ dysgalactiae} subspecies \emph{equisimilis} (SDE).
The former is a pervasive human-specific pathogen, and the latter
an opportunistic human pathogen.  Parts of the genomic configuration, including
human pathogenicity, of SDE could have originated in those of SPY.  
Our genome-wide survey of gene transfer indicated that gene transfer via homologous
recombination occurred relatively from SPY to SDE.
Although we did not find enrichment of virulence genes in the homologous recombining gene
transfer, we found evidence that they transferred more often via
non-recombining fashion than non-virulence genes. 
A genome-wide resource of recombining and non-recombining gene transfer for the study of the
genomes is provided so that research community in pathogenic bacteria can take
advantage of it.

@@

Bacteria evolves new biological features through horizontally transferred genes
from other lineages as well as vertically transmitted mutations on genes within
the same lineage.  Mutations on a gene modifiy relatively small portions of the
gene, including substitutions, insertions, and deletions of deoxyribonucleic
acids. Horizontal gene transfers can reshape the organization of a bacterial
genome, which could revamp the biological function of the bacterial species.
Recipient bacteria can substitute transferred genes for homologous parts of
their genomes. They can also integrate transferred genes in their genomes
without replacement. These two processes are discriminated by calling the former
\textit{recombining gene transfer}, and the latter 
\textit{horizontal gene transfer} \citep{Ochman2001,Lawrence2009}.  
Despite the two distinct evolutionary processes in bacteria, reseraches on
statistical analyses of complete genome sequences in efforts of understanding
bacterial evolution have paid little attention to a combined approach to
illuminating disparity of the two gene transfer processes.

HERE

\textbf{Gene transfer between SPY and SDE:} Through virulence genes
bacteria can infect humans
with a variety of diseases: e.g., food poisoning by \emph{Eschericha
Coli}, strep throat by \emph{Streptococcus pyrogenes}, tuberculosis
by \emph{Mycobacterium tuberculosis}, cholera by \emph{Vibrio cholera},
anthrax by \emph{Bacillus anthracis}, pneumonia by \emph{Streptococcus
pneumoniae} and \emph{Chlamydophila pneumoniae}, and influenza by
\emph{Haemophilus influenzae}, to name a few. \emph{S.\ pyogenes}
(SPY) belonging to group A streptococci (GAS) primarily infect humans
in the throat and skin \citep{Broyles2009}. It has high prevalence
throughout the world causing diseases ranging from mild illnesses
such as pharyngitis to severe invasive streptococcal diseases including
necrotizing fasciitis and toxic shock syndrome \citep{Cunningham2000a}.
A closely related species called\emph{ S.\ dysgalactiae} ssp.\emph{\ 
equisimilis}
(SDE) belonging to group C or G streptococci (GCS/GGS) used to be
generally regarded as a veterinary pathogen \citep{Vandamme1996}.
Researchers have been confirming that GCS/GGS can cause serious and
life-threatening streptococcal diseases to humans traditionally associated
with GAS \citep[e.g.,][]{Brandt2009}. Because SDE can cohabit with
SPY at the same ecological niche of tissue sites, and both may exhibit
overlap in the disease profile, the close proximity
may allow the two species to interact with each other by exchanging
parts of their genomes. It has been suspected that opportunistic pathogenicity
of SDE may have been facilitated by the very genetic interaction between
the two species. 
Although virulence genes were suspected to be acquired via non-recombining gene
transfer rather than recombining transfer \citep{Groisman1994}, systematic
approaches to elucidating relative effects of the two types of gene transfer 
rarely have been applied. This unfortunate situation motivated us to study the
closely related bacteria as efforts of better understanding relative 
association of gene transfer with virulence genes.
We hoped that understanding their genetic exchanges
would enhance our knowledge of mechanisms of their pathogenicity on
humans.

\textbf{Previous studies on gene transfer and its directionality between
SPY and SDE:} 
Because of the close interaction between SPY and SDE, 
possible donor and recipient relationship of the two species were noted before.
However, previous researches in gene transfer between
the two pyogeneic species of SPY and SDE were limited to small portion
of the bacterial genome, mostly focusing on several genes 
\citep{Kalia2004,Towers2004,Bessen2005}.
For example, \citet{Sachse2002} showed gene transfer from SDE to
SPY for the downstream of the genomic region of streptococcal pyrogenic
exotoxins \emph{speG}. \citet{Davies2007} investigated phage 3396
from SDE as a vehicle of gene transfer from SDE to SPY. \citet{Davies2009}
demonstrated that an integrative conjugative element of SDE allowed
gene transfer from SDE to other streptococci. On the contrary, \citet{Kalia2001}
attempted to show that the direction of gene transfer from SPY to
SDE was more dominant than the other direction, albeit unfortunately
appeared to be premature. Because their researches on direction of
gene flow were limited to only small portion of regions in the genomes,
it remains to be elusive to determine the genome-wide degrees to which
genomic segments flow between the two species and directions of thereof.
Cite this: \citep{Ahmad2009}.

\textbf{Methods of studying gene transfer in bacteria:} 
Studies on horizontal gene transfer in prokaryotes have focused on genes in very
large sets of species \citep{Koonin2001} using a great deal of genomes
available.  These studies gave us insights into complex evolutionary histories
of bacteria and the Archaea.  As sequencing costs plummet \citep{Mardis2011},
bacterial genome sequencing produces genomes of high quality even within
bacterial species \citep{Tettelin2009a}.  With more individual genomes per
species we now can go beyond the broad study on gene transfer towards
fine-scaled history between very closely related prokaryotes.  For example,
\citet{Liu2009} studied gene transfer events in \emph{Lactobacillus bulgaricus}
and \emph{Streptococcus thermophilus} using the compositional approach developed
by \citet{Karlin2001}. \citet{Luo2011} investigated gene transfer events in
\emph{Eschericha Coli} using the embedded quartet decomposition analysis
developed by \citet{Zhaxybayeva2006}. \citet{Hamady2006} combined phylogenetic
and compositional approaches to detecting horizontal gene transfer.
\citet{Caro-Quintero2011} studied a number of \emph{Shewanella baltica} genomes
to identify the genetic elements that enabled the species to adapt to redox
gradients. While compositional approaches were effective in analyzing
genome-scale data, phylognetic approaches have allowed to understand
evolutionary forces of gene exchange between species
\citep[e.g.,][]{Bapteste2004,Simonson2005,Beiko2009} by comparing differences
between gene trees and a reference species tree.  However, many of the methods
did not have explicit models of homologous recombining gene transfer that was
one of important evolutionary processes in horizontal gene transfer.  
% Overall,
% the development of methods for downstream analyses using bacterial genomes
% appears to lag behind that of high-throughput sequencing technology
% \citep[e.g.,][]{Green2011}.  
Recently, a method of detecting recombination 
with whole bacterial genomes \citep{Didelot2010} was developed. The explicit
model of homologous recombination allowed them to study 
flows of genetic segment between \emph{Bacillus cereus} species in the core
genomes that shared by most of individuals. 
In addition to recombination acting on bacterial genomes, genome rearragement,
translocation, duplication, loss, and more complex processes would govern the
evolution of bacteria. These processes, especially gene duplication, loss, and
non-recombining gene transfer, would result in more complex relationship between
even closely related bacteria. There have been progresses in 
methods of inferring gene duplication, loss, and transfer
events for gene families \citep{Doyon2011}. 
We noted that
the first approach could be used to infer transfer of genomic segments
due to homologous recombination, and the second could do similarly
due to non-recombining gene transfer.

%\textbf{Last paragraph before methods:} 
With these
methods for detecting gene transfer we interrogated gene flows
between SPY and SDE by measuring the relative impact of
two different processes of gene transfer on the pathogenicity of SPY
and SDE. This was possible thanks to recent release of bacterial
genomes of SPY and SDE. 
We here provided evidence that non-recombining
gene transfer was more effective for acquisitions of virulence genes
than homologous recombining gene transfer in SPY and SDE. 
While recombination rates varied along
the genomes, a few hot spots of recombination were observed. We found
directional gene flows in homologous recombining gene transfer process from SPY
to SDE. 
We showcased gene fragments of the inferred homologous gene transfers with high
posterior probabilities, mostly appearing to move from SPY to SDE.

\section{Methods and Materials}

\subsection{Data preparation}
Two genomes of \emph{S.\ pyogenes} (SPY), and two of \emph{S.\ dysgalactiae}
ssp.\emph{\ equisimilis} (SDE) were downloaded from NCBI with the following 
accessions:
NC\_004070 {Beres2002}, NC\_008024 \citep{Beres2006}, NC\_012891
\citep{Shimomura2011}, and CP002215 \citep{Suzuki2011}. We refer to these 
samples as SPY1, SPY2, SDE1, and SDE2, respectively. One genome
of \emph{S.\ dysgalactiae} 
ssp.\emph{\ dysgalactiae} was provided
by \citet{Suzuki2011} with accession identifiers of SddyATCC27957, referred to 
here as SDD. We also downloaded the genome (NC\_012471) of 
\emph{S.\ equi} ssp.\emph{\ equi} strain 4047 referred to as SEE.

\subsection{ClonalOrigin}
We present a short summary of the statistical framework developed by
\citet{Didelot2007} and \citet{Didelot2010} for studying bacterial 
recombination. A detailed account was found in the two papers and in the user 
guide available at \url{http://code.google.com/p/clonalorigin}. 
The analysis required first identifying sufficiently long contiguous blocks of 
high confident alignment. Multiple alignments with the six genomes
led to a dramatic reduction in the core 
genome length compared with those without SEE genome.
Because blocks of long multiple sequence alignments
were needed in order to obtain reliable 
estimates of recombination rate, we omitted the SEE sample 
from the ClonalOrigin analysis. We used progressiveMauve v2.3.1 
\citep{Darling2004,Darling2010} with default options
to align the five genomes.
We then used stripSubsetLCBs, a program of the ClonalOrigin package
to identify 276 blocks of sequence alignments with a length threshold of 
1500 base pairs.  We excluded two of these
blocks because they contained long consecutive gaps. Using these blocks, we ran 
ClonalFrame v1.1 available at
\url{http://www.xavierdidelot.xtreemhost.com} to obtain an 
estimate of the species tree relating the five genomes. ClonalFrame also 
outputs estimates of additional parameters, such as the average recombinant 
tract length and the ratio between recombination and mutation rates, however, 
these were not used in the subsequent analysis. We used a single Markov chain 
Monte Carlo (MCMC) with 10000 burn-in generations and 10000 sampling 
generations to obtain the tree estimate with branch lengths, 
and seven additional independent runs to validate adequate 
convergence of the first.

The next step of the analysis involved running ClonalOrigin in order to infer 
recombination histories across the alignment blocks. ClonalOrigin performs MCMC 
sampling of population parameters and recombination histories, given the 
species tree. The population parameters include mutation rate, 
recombination rate, and average tract length of a recombinant edge. A 
recombination history is represented by a \emph{recombinant tree}, which is a 
species tree augmented with a set of \emph{recombinant edges}. 
Although \citet{Didelot2010} described the concept of a recombinant tree
in detail, we elected to use Figure \ref{fig:clonalorigin} 
to illustrate it in order to clarify
summarization procedures we employed of posterior samples of recombinant trees.
Each recombinant 
edge is defined by time and branch of origin, time and branch of destination, 
and genomic interval in which it affects. The branch of origin is referred 
to as the \emph{donor} and the destination branch is referred to as the 
\emph{recipient}. A recombinant tree can be viewed as a restricted version of 
an ancestral recombination graph (ARG), and as such it induces a local 
genealogy for each site in the genomic block it describes. ClonalOrigin could
jointly sample population parameters and 
recombinant trees according to the posterior probabilities defined by a model 
approximating the coalescence process with recombination.
% \texttt{MAYBE ELABORATE MORE ON RESTRICTIONS OF CLONAL-ORIGIN'S MODEL? 
% SC:I HAVE NOT THOUGHT ABOUT MORE RESTRICTIONS ON THE RECOMBINANT TREES.}

% ClonalOrigin is applied in two stages. In both stages, sampling is performed 
% separately for the different alignment blocks. In the first stage, full 
% sampling is performed to obtain block-specific estimates of the model 
% parameters integrating out the possible recombinant trees. The block-specific 
% estimates of mutation rate, recombination rate, and average tract length are 
% then combined using a weighted median scheme to obtain a global estimate for 
% each parameter. In this scheme, each estimate is weighted according to the 
% length of the block with which it is associated, down-weighing the less 
% reliable estimates obtained from short blocks. In the second stage of 
% ClonalOrigin, the model parameters are fixed to their global estimate across 
% all blocks, and only the recombinant tree is sampled. Later this section we 
% describe several summaries of this posterior sample that are used in our 
% analysis. We used ClonalOrigin subversion r19 (downloaded from 
% \url{http://clonalorigin.googlecode.com/svn/trunk}). For the first stage, we 
% used two independent chains with 1,000,000 burn-in generations and  10,000,000 
% sampling generations, sub-sampled every 100,000 generations. The first chain 
% was used to obtain the block-specific estimates, and the second chain was used 
% to validate adequate convergence of the first. For the second stage, we used a 
% single chain with 10,000,000 burn-in generations and  100,000,000 sampling 
% generations, sub-sampled every 100,000 generations.

Two stages of MCMC with ClonalOrigin
were employed as described in \citet{Didelot2010}. 
Both stages of MCMC were performed independently for each of the 274 alignment blocks. 
We used ClonalOrigin subversion r19 that had been downloaded from 
\url{http://clonalorigin.googlecode.com/svn/trunk}. 
In the first MCMC we obtained estimates of 
mutation rate, recombination rate, and recombinant tract length. 
Two independent chains were executed with $10^6$ burn-in generations and $10^7$ 
sampling generations, sub-sampled every $10^5$-th generations. One of the two chains
was used to obtain the block-wise mean estimates, and the other chain was used 
to validate convergence of the first stage of MCMC. 
The estimators of the three population parameters were medians weighted over
block lengths, which down-weighed less reliable estimates obtained from shorter
alignment blocks. 
The second stage of MCMC with ClonalOrigin was executed with the three population
parameters fixed to the estimated values in the first stage. 
Two independent chains of the second stage
of ClonalOrigin's MCMC were also executed with a total generation being
$1.1\times10^8$. The first $10^7$ generations of each chain were discarded
as burn-in. The remaining $10^8$ generations were sub-sampled
every $10^5$-th generation.
We now describe the method of summarizing
the posterior sample to examine aspects of recombinant trees from the second
stage of ClonalOrigin's MCMC.

\citet{Didelot2010} presented two procedures of summarizing recombinant trees by
viewing trees from two different angles: one from recombinant trees by
aggregating all of the sites, and another from site-wise recombinant edge
changes to the species tree. By following their approaches
we quantified more aspects of recombinant trees: from the view of recombinant
trees with aggregated sites we counted gene tree topologies, and from the view of
site-wise recombinant edge changes we quantified recombination intensity along
the genome. 
We first describe procedures of summarizing recombinant trees from
the first viewpoint.
To summarize the posterior samples of ClonalOrigin we examined
the distribution of topologies of local genealogies induced by the sampled 
recombinant trees. Recall that a recombinant tree induces a local genealogy for 
each site along the analyzed block. Figure \ref{fig:clonalorigin} shows six
local gene trees including one that is the same as the species tree from the
most left-hand side of the block of 100 base pairs to the right.
We expected the dominant topology across 
sampled genealogies to be concordant with that of the species tree. However, 
other frequently sampled topologies were expected to be potentially informative about 
common recombination events. There are 105 possible 
rooted topologies relating five taxa, which allowed us to count topologies
induced by a posterior sample of recombinant trees.
Another useful summary suggested by 
\citet{Didelot2010} was to classify recombinant edges by types defined according 
to donor/recipient branch pairs, and record the number of edges sampled for
each recombinant edge type. For instance, our species tree has nine branches
including rooting branch and 
81 branch pairs, out of which 60 are possible donor/recipient pairs 
by prohibiting recombinant edges from going backward in time. The posterior 
expected number of edges of each type can be compared to the prior expected 
number determined by the three population parameters fixed in the second stage of 
ClonalOrigin. This comparison was displayed in a heat map of 
Figure \ref{fig:Heatmap-of-recombination}, which indicated branch pairs that 
might experience an excess or depletion of recombination events across all
blocks.

For the second viewpoint of recombinant trees
\citet{Didelot2010} suggested an approach used to trace
recombination intensity along the genome. 
For each edge type and each site, we recorded the sampling frequency of a 
recombinant edge of the type at that site (e.g., Figure S\ref{fig:muts}).
Figure \ref{fig:clonalorigin}B illustrates the posterior probability that a site
experiences influx from a donor branch using a single recombinant tree.
% This sampling frequency is an 
% estimate of the posterior probability taking into consideration the 
% sequence data and the setting chosen for the model parameters in the second 
% stage of ClonalOrigin. 
Summing these posteriors over a certain set of edge 
types, we could obtain a measure of recombination intensity along the genome 
(Figure \ref{fig:clonalorigin}B). 
This allowed us to find genomic regions enriched for recombination events in general, 
or regions enriched for recombination events of certain types (e.g.,
between SPY and SDE).  We note that the recombination intensity measure is not 
normalized to the interval $[0,1]$. Rather giving probabilistic meanings to the
recombination intensity we interpreted it as an average number of recombinant
edge types per site.
The measure appeared to act as an 
effective indicator of recombination intensity 
(see section of \textbf{Simulation Study}). 
Our simulation study indicated that edges of certain types were better indicators of 
true recombination events than others. For instance, recombinant edges which 
resulted in a recombinant tree with the same topology as the species tree appeared to 
be more prone to spurious sampling. 
Such edges were ones where the donor was the 
same branch, sister branch, or parent branch of the recipient branch. 
Therefore, in some cases we elected to remove such edges from consideration 
when measuring recombination intensity.
\texttt{IG: WE STILL NEED TO CHECK AND REFINE THIS IDEA. SC: WORKS ARE BEING DONE
FOR THIS ISSUE OF REMOVING UNINFORMATIVE RECOMBINANT EDGES}.

\subsection{Mowgli}
While ClonalOrigin could be used to understand recombining gene
transfers within the conserved core genome, we sought to understand
the history of the remainder of the genome where more complex
evolutionary processes occur, such as gene duplications, losses, and
horizontal transfers.  In recent years, several methods have been
developed for inferring these evolutionary events in gene families by
reconciling gene trees with a species tree
\citep{David2011,Doyon2011,Tofigh2011}.  In this analysis, we
complemented our study of recombination by using Mowgli
\citep{Doyon2011} to characterize events that can change the copy
number of a gene, such as duplications, losses, and non-recombining
horizontal gene transfers.  In contrast to ClonalOrigin, Mowgli
assumes a parsimony model for inferring events and it uses
pre-determined gene clusters where genes are treated as atomic units.
%The advantage of using Mowgli was that it did not assume the sequence
%data contains a single copy of each homologous segment, and it
%attempted to explain different copy numbers by a series of gain and
%loss events.  We here were interested in the approach of Mowgli to
%study non-recombining gene transfers.

We obtained gene family clusters defined in \citet{Suzuki2011} 
using the OrthoMCL method \citep{Li2003} over the full set of six genomes
including \emph{S.\ equi} subsp.\emph{\ equi} strain 4047 (SEE). 
For each family cluster,
the corresponding protein sequences were aligned using the MUSCLE
program \citep{Edgar2004a} and nucleotide alignments were constructed
by mapping nucleotide sequences onto each protein sequence alignment. We then
identified recombination within gene alignments using the Single
Breakpoint Recombination (SBP) method from the HyPhy package
\citep{KosakovskyPond2006}.  Using the Akaike Information Criterion (AIC), SBP
inferred that 47.5\% (1,094) of gene families contained at least one
topology-changing recombination. Each of these families was then
split into two gene fragments.
Gene trees were constructed from nucleotide alignments for each
gene fragment using the RAxML program \citep{Stamatakis2006}, and were
rooted by minimizing inferred duplications and losses.  Each rooted
gene tree was then reconciled using the Mowgli program
\citep{Doyon2011}, which inferred the most parsimonious gene
duplication, loss, and transfer events for each family.  In the case
of multiple maximum parsimonious reconciliations, one was chosen
uniformly at random.  Unless specified otherwise, default setings were used for
all programs.
Since the method of Mowgli did not directly model recombination events, 
recombination-based transfers were inferred in a post-processing step, by 
checking whether the transfer was coupled with a loss of a homologous gene in 
the recipient branch (indicating gene replacement rather than gain). 
A transfer event was classified as non-recombining if the 
destination species of that transfer had an extant descendant that contained
copies of that genes which were descendants of the transferred gene as well as 
copies which were non-descendants.
We noted that this parsimonious approach was imperfect and would tend to 
undercount events, especially recombining gene transfer events, which were
modeled using two events: a transfer and a loss. Nevertheless, this analysis 
allowed us to study non-recombining gene transfers, which were not modeled by 
ClonalOrigin, and to validate some of the ClonalOrigin results on recombining 
gene transfer.

\subsection{Functional categories association}
We performed a series of analyses to check for functional gene categories 
enriched for certain types of gene transfer events. Functional categories were 
assigned in the same manner as \citet{Suzuki2011}. In short, blastP was used to 
compare Streptococcus genes to bacterial proteins from the Uniref90 database.  
Matches with E-values less than $1.0\times10^{-5}$ were assumed to have the same Gene Ontology (GO) 
classification as the target gene, as determined from the uniProt GOA database. 
We also classified a category of  \emph{virulent genes} using the virulence 
profile described in \citet{Suzuki2011}. As virulent genes we used 207 loci, in 
which at least one of the genes in the gene family was identified by that 
profile.
Each gene covered by our analysis was associated with a recombination intensity 
score from the ClonalOrigin analysis and scores from the Mowgli 
analysis corresponding to number of duplications, losses, non-recombining 
gene transfers, and homologous recombining gene transfer events. 
\texttt{The recombination intensity scores were taken to be average site-wise intensity 
scores. SC: THIS NEEDS TO BE REANALYZED WITH RECOMBINATION INTENSITY WITH ONLY
TREE-TOPOLOGY CHANGING RECOMBINANT EDGES.} 
Mann-Whitney tests were used to identify GO categories 
whose genes showed significantly elevated values in each score. 
A 5\% False Discovery Rate (FDR) correction \citep{Benjamini1995} was used to 
determine an appropriate significance threshold for our p-values.

\subsection{Simulation Study}
We conducted experiments on simulation data to check the accuracy of 
estimates obtained by ClonalOrigin in a scenario similar to ours. We used a 
simulation feature of ClonalOrigin to simulate the process of mutation and 
recombination across the five species.
We started by testing the accuracy of the global estimates obtained from the 
first stage of ClonalOrigin for the three population parameters.
Ten recombinant trees were sampled given the tree parameters and the species
tree  set to the values inferred from the 
five bacterial genomes (see section \textbf{Results}).
Each of the recombinant trees was used to simulate
the Jukes-Cantor model for DNA sequence evolution to create 274 blocks of
alignments. 
The block lengths configuration was the same as that of the blocks 
used in the analysis of the five genomes.
We ran ClonalOrigin 
on the ten replicate data sets, and obtained global estimates by applying the 
weighted median scheme for the block-specific estimates. The means of the ten 
global estimates of mutation rate per site and recombination rate per site were 
0.081 with standard deviation (SD) of $0.0011$, and 0.011 with SD of $0.0003$, 
respectively, centering
around the true values: 0.081 and 0.012, respectively. The mean 
of the ten estimates of the recombinant tract length was 919 bp (SD 28), which 
was significantly higher than the true value of 744 bp. This was not entirely 
surprising as recombinant tract lengths were known to be difficult to 
estimate from relatively short alignment blocks \citep{Didelot2010}. When we 
performed similar simulations which much longer blocks (10000 bp), 
ClonalOrigin was able to estimate the average recombinant tract length 
accurately (data not shown).

% \texttt{IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.}
Next we evaluated the accuracy of statistics estimated in 
the second stage of ClonalOrigin. We used 100 replicates of data sets obtained
by first
sampling ten recombinant trees given the mutation rate, recombination rate, and
tract length, and generating ten independent 
simulations of sequence evolution using each of the recombinant trees. For each measure 
of interest we thus obtained 100 separate estimates, which we could compare 
with either
the expected value from the ten recombinant trees, or 
true value from one of the recombinant trees. 
First, we tested the number of 
recombinant edges of each of the sixty different types. Figure S\ref{fig:h1} 
plots  the distribution of observed counts against the 
respective expected count given the model parameters and species tree 
for each recombinant edge type.
\texttt{FIX ``TRUE" TO BE EXPECTED VALUE RATHER THAN AVERAGE OVER THE 10 REC. 
TREES. SC: OKAY. LET'S FIND THE PRIOR EXPECTED NUMBER OF EVENTS. THIS MUST BE
GIVEN SOMEWHERE}. 
The estimated numbers appeared to be slightly over-estimated, but in 
very good correlation with the expected numbers.

With the slight over-estimation of recombination events we were concerned with 
any potential biases hidden in ClonalOrigin implementation. Because the 
simulated data were generated with no directionality between SPY and SDE,  if 
we estimate recombination events from the simulated data sets using 
ClonalOrigin and compare those estimates from those from the real data set, 
this could remove biases in ClonalOrigin implementation. We could use inferred 
estimates of recombination from the simulated data under the prior model of 
ClonalOrigin as the prior expected number of recombination. Figure 
S\ref{fig:h2} shows a plot of numbers of recombination events from the real 
data against those from the simulated data sets. Note that the numbers of 
recombination events from simulated data sets appear to deviate from the line 
of slope being 1 and intercept being 0. The numbers of recombination events for 
the recombinant edges from SPY to SDE increase by more than 2.5 fold whereas 
those for the recombinant edges from SDE to SPY increase by about 1.7 fold.  
This was not surprising because inference of number of recombinant edges was 
quite accurate. We considered that subtraction of effects of number of 
recombinant edges inferred from simulated data and subtraction of effects of 
the prior expected number of recombinant edges were effectively equivalent. 
However, this comparison between inferrence from simulated data and that of 
real data set suggested that the directional difference between SPY and SDE was 
due to neither lack of power of ClonalOrigin nor hidden bias thereof.

%\textbf{Simulation study for checking the surrogate measure of recombination intensity:} 
We tested if our measure of recombination intensity provided a reliable 
estimate of the true intensity. For this purpose, we artificially annotaded our 
synthetic blocks to genes according to the real annotations along the real 
bacterial alignments. For each gene in each of the 100 data sets, we plotted 
the average estimated recombination intensity along that gene against the true 
average intensity, i.e., the average number of different recombinant edge types 
implied by the recombinant tree used in simulation (Figure S\ref{fig:ri1}). 
Although the scatter plot did not strictly align along the $X=Y$ line, there 
was clear correlation of the estimated and true values (Pearson's correlation 
coefficient of 0.63). Estimated values appeared to be driven somewhat by the 
prior implied by the population parameter setting, explaining why lower values 
were over-estimated and higher values were under-estimated. Nonetheless, the 
correlation indicated that the measure of recombination intensity used in our 
functional gene category served as an adequate indicator of actual intensity of 
recombination events along genes.

@@

\texttt{IG: THIS SECTION REQUIRES ADDITIONAL WORK, PENDING INTERPRETATION AND 
REFINEMENT OF RESULTS. }

\texttt{IG: THESE RESULTS MIGHT SLIGHTLY CHANGE BY REMOVING CERTAIN EDGE TYPES 
FROM CONSIDERATION. SC: ANALYSES WILL BE DONE BY REMOVING SOME EDGES.}

@@

% simulation for ClonalOrigin stage 1
We hoped that we could understand the power
and limitations of ClonalOrigin under the parameter setup that we
inferred from our data set.

>>>>

Inferring homologous recombination and horizontal gene transfer 
using whole genomes of \textit{Streptococcus pyogenes} 
and\textit{ S.\ dysgalactiae} subspecies \textit{equisimilis}

@@

phosphotransferase system (PTS), ATP-binding cassette (ABC), Aminoacyl
tRNA synthetase (aaRS), ancestral recombination graph (ARG), group
A streptococci (GAS), group C streptococci (GCS), group G streptococci
(GGS), \textit{Streptococcus pyogenes} (SPY), \textit{S.\ dysgalactiae}
subspecies \textit{dysgalactiae} (SDD), \textit{S.\ dysgalactiae} subspecies
\textit{equisimilis} (SDE).

@@

\section*{Abstract}
Homologous recombination and horizontal gene transfer are investigated with
complete genomes of closely related bacterial species of \textit{Streptococcus
pyogenes} (SPY) and \textit{S.\ dysgalactiae} subspecies \textit{equisimilis}
(SDE).  The former is a pervasive human-specific pathogen, and the latter an
opportunistic human pathogen.  Parts of the genomic configuration, including
human pathogenicity, of SDE could have originated in those of SPY.  Our
genome-wide survey of gene transfer indicated that gene transfer via homologous
recombination occurred relatively from SPY to SDE.  Although we did not find
enrichment of virulence genes in the homologous recombining gene transfer, we
found evidence that they transferred more often via non-recombining fashion than
non-virulence genes.  A genome-wide resource of recombining and non-recombining
gene transfer for the study of the genomes is provided so that research
community in pathogenic bacteria can take advantage of it.

@@

Bacteria infect humans with a variety of diseases: e.g., food poisoning
(\textit{Eschericha Coli}), strep throat (\textit{Streptococcus pyrogenes}),
tuberculosis (\textit{Mycobacterium tuberculosis}), cholera (\textit{Vibrio
cholera}), anthrax (\textit{Bacillus anthracis}), pneumonia
(\textit{Chlamydophila pneumoniae}), and influenza (\textit{Haemophilus
influenzae}), to name a few.  \textit{Streptococcus pyogenes} (SPY) belonging to
group A streptococci (GAS) primarily infect humans in the throat and skin
\citep{Broyles2009}. It has high prevalence throughout the world causing
diseases ranging from mild illnesses such as pharyngitis to severe invasive
streptococcal diseases including necrotizing fasciitis and toxic shock syndrome
\citep{Cunningham2000a}.  Although \textit{Streptococcus dysgalactiae}
subspecies \textit{equisimilis} (SDE) belonging to group C or G streptococci
(GCS/GGS) used to be generally regarded as a veterinary pathogen
\citep{Vandamme1996}, it has been reported to cause serious and life-threatening
streptococcal diseases to humans traditionally associated with GAS
\citep[e.g.,][]{Brandt2009}.  Because of overlap in the disease profile and the
ecological niche of tissue sites betwen SPY and SDE, these bacteria may as well
interact with each other by exchanging parts of their genomes.

The research topic on gene exchanges between SPY and SDE has been discussed in
various ways. \citet{Sachse2002} showed gene transfer from SDE to SPY for the
downstream of the genomic region of streptococcal pyrogenic exotoxins
\textit{speG}. \citet{Davies2007} investigated phage 3396 from SDE as a vehicle
of gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that an
integrative conjugative element of SDE allowed gene transfer from SDE to other
streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
directionality of housekeeping gene transfer between SPY and SDE.  Notably,
\citet{Kalia2001} attempted to show that the direction of gene transfer from SPY
to SDE was more dominant than the other direction.  Because these researches on
the controversial issue, directionality of gene transfer, have been limited to
relatively small regions in the genomes, it remains to be elusive to determine
the genome-wide degrees to which genomic segments flow between the two species,
and directions of thereof.

As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
produces individual genomes of high quality even within a bacterial species
\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
used genes in very large sets of species \citep{Koonin2001} from a great deal
of genomes available.  Although this type of study gave us insights into complex
evolutionary histories of prokaryotes, with multiple individual genomes per
species we now can go beyond the broad study on gene transfer towards
fine-scaled history of gene transfer between very closely related prokaryotes.
For instance, \citet{Liu2009} studied gene transfer events in
\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
the compositional approach developed by \citet{Karlin2001}. \citet{Luo2011}
investigated gene transfer events in \textit{Eschericha Coli} using the embedded
quartet decomposition analysis developed by \citet{Zhaxybayeva2006}.
\citet{Hamady2006} combined phylogenetic and compositional approaches to
detecting horizontal gene transfer.  \citet{Caro-Quintero2011} studied a number
of \textit{Shewanella baltica} genomes to identify the genetic elements that
enabled the species to adapt to redox gradients.

Bacteria evolves new biological features through horizontally transferred genes
from other lineages and vertically transmitted mutations on genes within
the same lineage.  Mutations on a gene modifiy relatively small portions of the
gene, including substitutions, insertions, and deletions of deoxyribonucleic
acids. Horizontal gene transfers can reshape the organization of a bacterial
genome, which could revamp biological functions of the bacterial species.
Recipient bacteria can substitute transferred genes for homologous parts of
their genomes. They can also integrate transferred genes into their genomes
without replacement. These two processes are here discriminated by calling the
former \textit{recombining gene transfer}, and the latter \textit{horizontal
gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the two distinct
evolutionary processes in bacteria, reseraches on statistical analyses of
complete genome sequences in efforts of better understanding bacterial evolution
have paid little attention to a combined approach to illuminating disparity of
the two gene transfer processes.

Homologous recombining gene transfer could be studied more readily using the
core genomes from multiple individual bacterial genomes while horizontal gene
transfer might be more easily detected using the pan genomes. Recently, a method
of inferring homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segment between bacterial individuals of \textit{Bacillus
cereus} species in the core genomes.  More complex evolutionary processes acting
on bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  Because of the complexity, these processes do not
readily lend themselves to statistically more sound approaches such as maximum
likelihood. Instead, a parsimony-based approach to inferring gene gain, loss,
and transfer events was developed by  \citep{Doyon2011}.  We noted that the
first approach could be used to infer transfer of genomic segments due to
homologous recombination, and the second could do similarly due to horizontal
gene transfer. We expect that model-based approaches using whole bacterial
genomes will be developed in the near future.

The complete genomes of the two spcies of SPY and SDE presented a unique
opportunity to study gene transfer over relatively recent time scales.  We used
both of the model-based approach for inferring homologous recombination and the
parsimony-based one for detecting gene gain, loss, and horizontal transfer. We
also compared both of the results in recombining gene transfer events to show
that the parimony-based method inferred gene transfer events in similar
biological functional categories to those inferred with the model-based
approach. This is the first attempt to analyaze SPY and SDE in genome-wide scale
for detecting gene transfer events not only in the core genome but also in the
pan genome, which make the analysis complete in terms of the amount of used
genome data. Our analysis of core genomes also showed for the first time that
the net directionality of, at least, recombining gene transfer was more from SPY
to SDE.

@@

@@

% simulation for ClonalOrigin stage 1
We hoped that we could understand the power
and limitations of \texttt{ClonalOrigin} under the parameter setup that we
inferred from our data set.
